A Data Restore Model\\ for Reproducibility in Computational Statistics

### 2013-05-07 paper-data-restore-model.tex
language

If implemented in big scale as a holistic information infrastructure for statistical research data management, the presented approach has further implications on workflows and gives ideas for a big picture.
Data archives could offer the download of data templates along with the actual data, so that researchers have the precise references right from the beginning of their research.
Alternatively, researchers could pull the data on the basis of such data template from within their research tool, provided that a respective plugin is available.
The advantage in the latter case is that this procedure could be embedded within their program code, so that no further modifications of data and code need to be done at the time of publishing.
The researcher simply submits code and data templates used, and other researchers can rerun the same analysis without any manual arrangements or collection of resources, simply by executing the code.

In case the researcher used own data, the data would have to be submitted to an affiliated archive\footnote{Researchers have to archive their data, either locally or at another archive. By "affiliated archive" we denote this repository.} first where the respective data template and its IDs involved are generated first.
The archive returns the template to the researcher who can then publish it in the same way as described above.
This procedure must be supported with tools for mapping and generation of IDs.
Furthermore, these tools should provide for further semantic annotation of the content (using the RDF Data Cube Vocabulary for example) so that documentation about the data is available as well.
Figure \ref{fig:modelprocessing} outlines the overall process.

@@

We introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
In addition, we also declare Web Service addresses for every namespace used (which is only one for our example) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
%At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.

@@

\begin{figure*}[htb]
\centering
\includegraphics[width=.95\textwidth]{img/datatemplate.png}
\caption{
%Excerpt from one of the data templates
%Data template excerpt from one of the data templates
Data template for one of the two data sets \textit{rk.raw}
}
\label{fig:datatemplate}
\end{figure*}

@@

We validate this experimental implementation with respect to the requirements developed in Section \ref{sec:req}.
As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).
The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can be done in one place. 
Since the use of RDF allows us to add an additional link between ID and value, we can separate the two and disclose all information about the data in very detail in machine-understandable form, and yet protect the values (\textit{Protection}).
Furthermore, when every value can be referred to from anywhere, a simple technique can be applied to enable format-independent reconstruction of data files for which we introduced the concept of a \textit{data template}.
Particularly the document-specific declaration of tokens enables a format-independent solution (\textit{Formats}) as this flexibility allows us to avoid syntactical clashes with the particular data format used.
We have used SPARQL endpoints as a generic implementation of the \textit{API} which only served for the purpose of data retrieval.
For actual application, a more elaborated API should be developed that could be based on Web Services.

>>>>

If implemented in big scale as a holistic information infrastructure for research data management in statistics, the presented approach has further implications on workflows and gives ideas for a big picture.
Data archives could offer the download of data templates along with the actual data, so that researchers have the precise references right from the beginning of their research.
Alternatively, researchers could pull the data on the basis of these data templates alone from within their research tool, provided that a respective plugin is available.
The advantage in the latter case is that this procedure could be embedded within their program code and no further modifications on data and code need to be done at the time of publishing.
The researcher simply submits code and data templates, and other researchers can rerun the same analysis without any manual arrangements or collection of resources, simply by executing the code.

In case the researcher used own data, the data would have to be submitted to an affiliated archive\footnote{Researchers have to archive their data, either locally or at another archive. By "affiliated archive" we denote this repository.} in beforehand where the respective data template and IDs involved are generated first.
The archive returns the template to the researcher who can then publish it in the same way as described above.
This procedure must be supported with tools for mapping and ID generation.
Furthermore, these tools should provide for further semantic annotation of the content (using the RDF Data Cube Vocabulary for example) so that documentation about the data is available as well.
Figure \ref{fig:modelprocessing} outlines the overall process.

@@

%We introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
In addition, we also declare Web Service addresses for every namespace used (which is only one for our example) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
%At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.

@@

\begin{figure*}[htb]
\centering
\includegraphics[width=.95\textwidth]{img/datatemplate.png}
\caption{
%Excerpt from one of the data templates
%Data template excerpt from one of the data templates
Data template for one of the two data sets (\textit{rk.raw})
}
\label{fig:datatemplate}
\end{figure*}

@@

We validate this experimental implementation with respect to the requirements developed in Section \ref{sec:req}.
As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Traceability}).
The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can be done in one place. 
Since the use of RDF allows us to add an additional link between ID and value, we can separate the two and disclose all information about the data in very detail in machine-understandable form, and yet protect the values (\textit{Protection}).
Furthermore, when every value can be referred to from anywhere, a simple technique can be applied to enable format-independent reconstruction of data files for which we introduced the concept of a \textit{data template}.
Particularly the document-specific declaration of tokens enables a format-independent solution (\textit{Formats}) as this flexibility allows us to avoid syntactical clashes with the particular data format used.
We have used SPARQL endpoints as a generic implementation of the \textit{API} which only served for the purpose of data retrieval.
For actual application, a more elaborated API should be developed that could be based on Web Services.
